Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2021-10-14 manuscript/manuscript.Rmd
convert percentage numbers to decimals

Model-based forecasts, which have played an important role in shaping public policy throughout the Covid-19 pandemic, represent an implicit combination of model assumptions and the researcher’s subjective opinion. This work explicitly separates human opinion and model-derived insights to assess and compare the relative strengths and weaknesses of both approaches. We compared purely opinion-derived forecasts of cases and deaths from COVID-19 in Germany and Poland, elicited from researchers and volunteers (“crowd forecasts”), against predictions from two semi-mechanistic epidemiological models. In addition, we compared these forecasts against an ensemble of model-based, but expert-tuned, forecasts, submitted to the German and Polish Forecast Hub by other research institutions (“Hub ensemble”). We found an ensemble of crowd-sourced predictions to outperform all other methods when predicting cases (Weighted Interval Score relative to the Hub ensemble at a two-week horizon: 89%), but not when predicting deaths (rel. WIS 126%). Crowd forecasts were severely overconfident (55% and 75% coverage of the 95% prediction intervals for cases and deaths, respectively) when compared to model-based predictions. Performance of the untuned models deteriorated quickly with increasing forecast horizon when assumptions were no longer met (rel. WIS > 164% for all four week ahead predictions).

@@

For cases, crowd forecasts had a lower mean weighted interval score (WIS, lower values indicate better performance) than both the renewal model and the Hub ensemble across all forecast horizons (Figure \@ref(fig:agg-performance-all)A) and locations (Figure \@ref(fig:performance-locations-rel)A). For two week ahead forecasts, mean WIS relative to the Hub ensemble (= 100%) was 89% for crowd forecasts and 140% for the renewal model (Table \@ref(tab:score-table-2)). Across all forecasting approaches, locations and forecast horizons, the distribution of WIS values was right-skewed, and average performance was heavily influenced by outliers (Figure \@ref(fig:distribution-scores)). Overall, low variance in forecast performance was closely linked with good mean performance (Figures \@ref(fig:agg-performance-all)H and and \@ref(fig:agg-performance-all)A), suggesting that the ability to avoid large errors was an important factor in determining overall performance. The impact of outlier values was especially pronounced for the renewal model, which had more outliers (Figure \@ref(fig:distribution-scores), as well as the highest standard deviation of WIS values (relative WIS sd 1.54 at the two weeks ahead horizon), while the ensemble of crowd forecasts  (rel. WIS sd 0.76) and the Hub ensemble (= 1) showed more stable performance.

>>>>

Model-based forecasts, which have played an important role in shaping public policy throughout the Covid-19 pandemic, represent an implicit combination of model assumptions and the researcher’s subjective opinion. This work explicitly separates human opinion and model-derived insights to assess and compare the relative strengths and weaknesses of both approaches. We compared purely opinion-derived forecasts of cases and deaths from COVID-19 in Germany and Poland, elicited from researchers and volunteers (“crowd forecasts”), against predictions from two semi-mechanistic epidemiological models. In addition, we compared these forecasts against an ensemble of model-based, but expert-tuned, forecasts, submitted to the German and Polish Forecast Hub by other research institutions (“Hub ensemble”). We found an ensemble of crowd-sourced predictions to outperform all other methods when predicting cases (Weighted Interval Score relative to the Hub ensemble at a two-week horizon: 0.89), but not when predicting deaths (rel. WIS 1.26). Crowd forecasts were severely overconfident (55% and 75% coverage of the 90% prediction intervals for cases and deaths, respectively) when compared to model-based predictions. Performance of the untuned models deteriorated quickly with increasing forecast horizon when assumptions were no longer met (rel. WIS > 1.64 for all four week ahead predictions).

@@

For cases, crowd forecasts had a lower mean weighted interval score (WIS, lower values indicate better performance) than both the renewal model and the Hub ensemble across all forecast horizons (Figure \@ref(fig:agg-performance-all)A) and locations (Figure \@ref(fig:performance-locations-rel)A). For two week ahead forecasts, mean WIS relative to the Hub ensemble (= 1) was 0.89 for crowd forecasts and 1.40 for the renewal model (Table \@ref(tab:score-table-2)). Across all forecasting approaches, locations and forecast horizons, the distribution of WIS values was right-skewed, and average performance was heavily influenced by outliers (Figure \@ref(fig:distribution-scores)). Overall, low variance in forecast performance was closely linked with good mean performance (Figures \@ref(fig:agg-performance-all)H and and \@ref(fig:agg-performance-all)A), suggesting that the ability to avoid large errors was an important factor in determining overall performance. The impact of outlier values was especially pronounced for the renewal model, which had more outliers (Figure \@ref(fig:distribution-scores), as well as the highest standard deviation of WIS values (relative WIS sd 1.54 at the two weeks ahead horizon), while the ensemble of crowd forecasts  (rel. WIS sd 0.76) and the Hub ensemble (= 1) showed more stable performance.
